---
title: "XCT results"
author: "Erin C Rooney"
date: "10/26/2020"
output: github_document
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE,
                      comment = "#>", 
                      fig.path = "images/"
                      )

```

## Measuring the Soil Pore Network response to freeze/thaw in permafrost soil aggregates using X-Ray Computed Tomography


## Pore Throat Diameters

<details>
  <summary>click to open stats</summary>

### Statistics
```{r Load}
breadthdata_csv = read.csv("processed/ftc_porethroatdist_july312020_2.csv") 
levels(as.factor(breadthdata_csv$trmt)) 
library(agricolae)
library(tidyverse)
library(PairedData)
library(ggpubr)
library(ggthemes)

```


```{r Statistics1}
breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(trmt = recode(trmt, "before " = "before"))
tool = breadthdata_csv %>% 
  filter(site=="tool")
before <- subset(tool, trmt == "before", breadth_dist, drop = TRUE)
after <- subset(tool, trmt == "after", breadth_dist, drop = TRUE)
pd <- paired(before, after)
d <- with(tool,
          breadth_dist[trmt == "before"] - breadth_dist[trmt == "after"])
shapiro.test(d)
res <- wilcox.test(breadth_dist ~ trmt, data = tool, paired = TRUE,
                   alternative = "less")
res
res$p.value

```

</details>

<details>
  <summary>click to open figures</summary>
  
### Figures
```{r Pore Throat Distributions}
library(PNWColors)
theme_er <- function() {  # this for all the elements common across plots
  theme_clean() %+replace%
    theme(legend.position = "top",
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 12),
          legend.key.size = unit(1.5, 'lines'),
          panel.border = element_rect(color="white",size=0, fill = NA),
          plot.title = element_text(hjust = 0.5, size = 14),
          plot.subtitle = element_text(hjust = 0.5, size = 12, lineheight = 1.5),
          axis.text = element_text(size = 12, color = "black"),
          axis.title = element_text(size = 12, face = "bold", color = "black"),
          
          # formatting for facets
          panel.background = element_blank(),
          strip.background = element_rect(colour="white", fill="white"), #facet formatting
          panel.spacing.x = unit(1.5, "lines"), #facet spacing for x axis
          panel.spacing.y = unit(1.5, "lines"), #facet spacing for x axis
          strip.text.x = element_text(size=12, face="bold"), #facet labels
          strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels
    )
}
theme_er2 <- function() {  # this for all the elements common across plots
  theme_bw() %+replace%
    theme(legend.position = "top",
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 12),
          legend.key.size = unit(1.5, 'lines'),
          panel.border = element_rect(color="black",size=2, fill = NA),
          
          plot.title = element_text(hjust = 0.5, size = 14),
          plot.subtitle = element_text(hjust = 0.5, size = 12, lineheight = 1.5),
          axis.text = element_text(size = 14, color = "black"),
          axis.title = element_text(size = 14, face = "bold", color = "black"),
          
          # formatting for facets
          panel.background = element_blank(),
          strip.background = element_rect(colour="white", fill="white"), #facet formatting
          panel.spacing.x = unit(1.5, "lines"), #facet spacing for x axis
          panel.spacing.y = unit(1.5, "lines"), #facet spacing for x axis
          strip.text.x = element_text(size=12, face="bold"), #facet labels
          strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels
    )
}

levels(as.factor(breadthdata_csv$sample))
str(breadthdata_csv)

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "40_50_16" = "Aggregate-1"))

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "40_50_28" = "Aggregate-2"))

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "28_38_12" = "Aggregate-3"))

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "28_38_28" = "Aggregate-4"))

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "41_50_16" = "Aggregate-5"))

breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(sample = recode(sample, "41_50_28" = "Aggregate-6"))
         
 
breadthdata_csv = 
  breadthdata_csv %>% 
  mutate(trmt = factor(trmt, levels = c("before", "after")))

tool = breadthdata_csv %>% 
  filter(site=="tool")

g = ggplot(tool, aes(x = breadth_um, y=breadth_dist, color = trmt))+
  geom_line(size = 1)+
  #geom_density(adjust=0.5)+
  
  labs (title = "Pore Throat Diameter Distribution",
        #subtitle = "After Freeze/Thaw",
        #caption = "Permafrost Soil Aggregate from Toolik, Alaska",
        #tag = "Figure 6",
        x = expression (bold ("Pore Throat Diameter, um")),
        y = expression (bold ("Distribution, %"))) 


g + theme_er() + 
  scale_color_manual(values = c("#b0986c", "#72e1e1"))+
  #scale_color_manual(values = pnw_palette("Anemone", 2, type = "discrete")) +
  #guides(fill = guide_legend(reverse = TRUE, title = NULL)) +
  facet_wrap(sample~.)

g + theme_er2() + 
  #scale_color_manual(values = c("#b0986c", "#72e1e1"))+
  scale_color_manual(values = pnw_palette("Bay", 2, type = "discrete")) +
  #guides(fill = guide_legend(reverse = TRUE, title = NULL)) +
  facet_wrap(sample~.)

g + theme_er2() + 
  #scale_color_manual(values = c("#b0986c", "#72e1e1"))+
  scale_color_manual(values = pnw_palette("Starfish", 2, type = "discrete")) +
  #guides(fill = guide_legend(reverse = TRUE, title = NULL)) +
  facet_wrap(sample~.)
```


</details>

## Pore Coordination Numbers

<details>
  <summary>click to open stats</summary>

### Statistics
```{r statistics2}
coornum_dat2 = read.csv("processed/porecoor_fixed.csv")
levels(as.factor(coornum_dat2$pore_coor))
coornum_dat2 = 
  coornum_dat2 %>% 
mutate(pore_coor = factor(pore_coor, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16")))
levels(as.factor(coornum_dat2$trmt))
tool = coornum_dat2[coornum_dat2$site=="tool",]
#Shapiro-Wilk normality test----------------------------
before <- subset(tool, trmt == "before", freq, drop = TRUE)
after <- subset(tool, trmt == "after", freq, drop = TRUE)
pd <- paired(before, after)
d <- with(tool,
          freq[trmt == "before"] - freq[trmt == "after"])
shapiro.test(d) # p value = 4.011e-11, not normally distributed
# Wilcoxon test Method 1
res <- wilcox.test(freq ~ trmt, data = tool, paired = TRUE)
res
res$p.value
```

</details>

<details>
  <summary>click to open figures</summary>

### Figures
```{r Pore Coordination Numbers}
ggplot(tool, aes(x = pore_coor, y = freq, fill = trmt)) +
  geom_boxplot() + 
  scale_y_continuous(labels = scales::percent) +
  theme_er() +
  scale_fill_manual(values = pnw_palette("Anemone", 2, type = "discrete")) +
  labs (title = "Pore Coordination Number Frequency",
        #caption = "Caption",
        #tag = "A",
        x = expression (bold ("Pore Coordination Number")),
        y = expression (bold ("Frequency, %"))) +
   guides(fill = guide_legend(reverse = TRUE, title = NULL))
```

</details>

## Air and Water-filled Pore Volumes

<details>
  <summary>click to open stats</summary>

### Statistics
```{r statistics3}
conn_csv = read.csv("processed/conn_unconn_aug52020.csv") 
level_order <- c('before', 'after')
tool = conn_csv %>% 
  filter(site=="tool")
conn.aov <- aov(conn_water_perc ~ trmt, data = tool)
summary.aov(conn.aov)
trmt_hsd = HSD.test(conn.aov, "trmt")
print(trmt_hsd)
conn_aov2 = aov(data = tool, unconn_pore_perc ~ trmt)
summary(conn_aov2)
conn_aov3 = aov(data = tool, conn_pore_perc ~ trmt)
summary(conn_aov3)
conn_aov4 = aov(data = tool, unconn_water_perc ~ trmt)
summary(conn_aov4)
```

</details>

<details>
  <summary>click to open figures</summary>

### Figures
```{r Pore volumes}
b1 = ggplot(tool, aes(x=factor(trmt,level_order), y=conn_water_perc, fill=trmt)) + geom_boxplot() +
  labs (title = "Connected Water-Filled Pores",
        # caption = "Permafrost Soil Aggregate from Toolik, Alaska",
        tag = "A",
        x = expression (bold (" ")),
        y = expression (bold ("Volume, %"))) +  scale_y_continuous(labels = scales::percent, limits = c(0, 0.09)) +
  theme_er() +
  scale_fill_manual(values = pnw_palette("Anemone", 2, type = "discrete")) +
  annotate("text", x = 1.5, y = 0.083, label = "p value < 0.05") +
  annotate("text", x = 1, y = 0.055, label = "A") +
  annotate("text", x = 2, y = 0.037, label = "B") 

  #guides(fill = guide_legend(reverse = TRUE, title = NULL))

b2 = ggplot(tool, aes(x=factor(trmt,level_order), y=conn_pore_perc, fill=trmt)) + geom_boxplot() +
  labs (title = "Connected Air-Filled Pores",
        # caption = "Permafrost Soil Aggregate from Toolik, Alaska",
        tag = "B",
        x = expression (bold (" ")),
        y = expression (bold ("Volume, %"))) +   scale_y_continuous(labels = scales::percent, limits = c(0, 0.09)) + 
  theme_er() +
  scale_fill_manual(values = pnw_palette("Anemone", 2, type = "discrete")) 
  # annotate("text", x = 2.25, y = 0.070, label = "P value < 0.5") +
  #guides(fill = guide_legend(reverse = TRUE, title = NULL))

b3 = ggplot(tool, aes(x=factor(trmt,level_order), y=unconn_water_perc, fill=trmt)) + geom_boxplot() +
  labs (title = "Unconnected Water-Filled Pores",
        #caption = "Permafrost Soil Aggregate from Toolik, Alaska",
        tag = "C",
        x = expression (bold (" ")),
        y = expression (bold ("Volume, %"))) +   scale_y_continuous(labels = scales::percent, limit = c(0, 0.02)) +
  theme_er() +
  scale_fill_manual(values = pnw_palette("Anemone", 2, type = "discrete")) 
  # annotate("text", x = 2.25, y = 0.070, label = "P value < 0.5") +
  #guides(fill = guide_legend(reverse = TRUE, title = NULL))

b4 = ggplot(tool, aes(x=factor(trmt,level_order), y=unconn_pore_perc, fill=trmt)) + geom_boxplot() +
  labs (title = "Unconnected Air-Filled Pores",
        #caption = "Permafrost Soil Aggregate from Toolik, Alaska",
        tag = "D",
        x = expression (bold (" ")),
        y = expression (bold ("Volume, %"))) +  scale_y_continuous(labels = scales::percent, limits = c(0, 0.02)) +
        theme_er() +
  scale_fill_manual(values = pnw_palette("Anemone", 2, type = "discrete")) 
  # annotate("text", x = 2.25, y = 0.070, label = "P value < 0.5") +
        #guides(fill = guide_legend(reverse = TRUE, title = NULL))

library(patchwork)
b1+b2+b3+b4+ #combines the two plots
  plot_layout(guides = "collect") # sets a common legend
```


</details>


### Session Info

<details>
  <summary>Session Info</summary>
  
```{r}
Sys.Date()
sessionInfo()
```

</details>
